---
title: "Stat 437 HW2"
author: 
     - Anthony Chelf (011743578)
header-includes:
   - \usepackage{bbm}
   - \usepackage{amssymb}
   - \usepackage{amsmath}
   - \usepackage{graphicx,float}
   - \usepackage{natbib}
output:
  html_document:
    df_print: paged
  pdf_document: default
fontsize: 11pt
always_allow_html: yes
---

```{r, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# General rule

Please show your work and submit your computer codes in order to get points. Providing correct answers without supporting details does not receive full credits. This HW covers:

- Advanced Visualizations via ggplot2: adjusting legends, fonts, orientation, and math expressions
- Visualizing networks as graphs
- Interactive visualization 

You DO NOT have to submit your HW answers using typesetting software. However, your answers must be legible for grading. Please upload your answers to the course space.

# Problem 1

Please refer to the NYC flight data `nycflights13` that has been discussed in the lecture notes and whose manual can be found at https://cran.r-project.org/web/packages/nycflights13/index.html. We will use `flights`, a tibble from `nycflights13`.

```{r, message=FALSE, warning=FALSE}
# packages to be used
library(dplyr)
library(ggplot2)
library(plotly)
library(nycflights13)
library(igraph)
library(igraphdata)
```
```{r}
# peaking at the nycflights13 data
head(flights)
```

You are interested in looking into the average `arr_delay` for 4 different `month` 12, 1, 7 and 8, for 3 different `carrier` "UA", "AA" and "DL", and for `distance` that are greater than 700 miles, since you suspect that colder months and longer distances may result in longer average arrival delays. Note that you need to extract observations from `flights`, and that you are required to use `dplyr` for this purpose.

```{r}
flights <- filter(flights, month %in% c(12,1,7,8),
                           carrier %in% c('UA', 'AA', 'DL'), distance > 700)
# change NA values for computing averages
flights = as.data.frame(flights)
flights[is.na(flights)] <- 0
```

The following tasks and questions are based on the extracted observations.

(1.a) For each combination of the values of `carrier` and `month`, obtain the average `arr_delay` and obtain the average `distance`. Plot the average `arr_delay` against the average `distance`, use `carrier` as facet; add a title "Base plot" and center the title in the plot. This will be your base plot, say, as object `p`. Show the plot `p`.

```{r}
# utilizes piping to find combination of values of month/carrier
# and computes averages of delay and distance
temp <- flights %>% select(month, carrier, arr_delay, distance) %>%
   filter(month %in% c(1, 7, 8, 12), carrier %in% c("AA", "UA", "DL"))
temp <- na.omit(temp)
temp <- temp %>% group_by(month, carrier) %>%
   summarise(mean_arr_delay = mean(arr_delay),
             mean_distance = mean(distance), .groups="keep") %>% as.data.frame()
# This portion is for adding the expression required in part 1b
mod_p = c(expression(alpha[1]), expression(beta[1,2]), expression(gamma^0))
temp$DF = factor(temp$carrier, labels = mod_p)
#generates the base plot
p <- ggplot(temp, aes(x = mean_arr_delay, y = mean_distance)) +
   geom_point() + facet_wrap(~carrier) + ggtitle("Base Plot") +
   theme(plot.title = element_text(hjust = 0.5))
# showing the plot p
p
```

(1.b) Modify `p` as follows to get a plot `p1`: connect the points for each `carrier` via one type of dashed line; code the 3 levels of `carrier` as $\alpha_1$, $\beta_{1,2}$ and $\gamma^{[0]}$, and display them in the strip texts; change the legend title into "My $\zeta$" (this legend is induced when you connect points for each `carrier` by a type of line), and put the legend in horizontal direction at the bottom of the plot; add a title "With math expressions" and center the title in the plot. Show the plot `p1`.

```{r}
# updates the base plot and adds the math expressions/dashed lines
p1 = p + geom_line(linetype = "dashed", aes(col=carrier)) + facet_wrap(~DF, labeller = label_parsed) + labs(col=expression(paste("My ", zeta, sep=""))) + scale_color_discrete(labels = mod_p) + theme(legend.position = "bottom", legend.direction="horizontal") + ggtitle("With Math Expressions")
# Show the plot `p1`.
p1
```


(1.c) Modify `p1` as follows to get a plot `p2`: set the font size of strip text to be 12 and rotate the strip texts counterclockwise by 15 degrees; set the font size of the x-axis text to be 10 and rotate the x-axis text clockwise by 30 degrees; set the x-axis label as "$\hat{\mu}$ for mean arrival delay"; add a title "With front and text adjustments" and center the title in the plot. Show the plot `p2`

```{r}
# creates an additional expression to be added along the x axis
mu_hat = expression(paste(hat(mu), " For Mean Arrival Delay", sep = ""))
# updates p1 to include a new title, and some rotations for various components
p2 = p1 + ggtitle("With Front and Text Adjustments") + theme(strip.text = element_text(size = 12, angle = 15), axis.text.x = element_text(size = 10, angle = -30)) + xlab(mu_hat)
# Show the plot `p2`.
p2
```

# Problem 2 

This problem requires you to visualize the binary relationship between members of a karate club as an undirected graph. Please install the R library `igraphdata`, from which you can obtain the data set `karate` and work on it. Create a graph for `karate`. Once you obtain the graph, you will see that each vertex is annotated by a number or letter. What do the numbers or letters refer to? Do you see subgraphs of the graph? If so, what do these subgraphs mean? 

```{r}
data("karate")
# creating the plot showing karate relationships
p3 <- simplify(karate, remove.multiple = T, remove.loops = T)
plot.igraph(p3, edge.arrow.size = 1, edge.color = "black")
```

A (John A) and H (Mr. Hi) represent who is running the karate class and the numbers refer to actors in the class. There are two subgraphs, one for node A and the other for node H.

# Problem 3

This problem requires to to create an interactive plot using `plotly`. If you want to display properly the plot in your HW answers, you may well need to set your HW document as an html file (instead of doc, docx or pdf file) when you compile your R codes.

Please use the `mpg` data set we have discussed in the lectures. Create an interactive, scatter plot between "highway miles per gallon" `hwy` (on the y-axis) and "engine displacement in litres" `displ` (on the x-axis) with the `color` aesthetic designated by "number of cylinders" `cyl`, and set the x-axis label as "highway miles per gallon" and y-axis label as "highway miles per gallon". You need to check the object type for `cyl` and set it correctly when creating the plot. Add the title "# of cylinders" to the legend and adjust the vertical position of the legend, if you can. For the last, you may look through https://plotly.com/r/legend/ for help.

```{r}
# creating scatter plot for mpg and updating the title for the legend/axis
p4 <- plot_ly(mpg, x = ~displ, y = ~hwy, color = ~cyl, type = "scatter", mode = "markers", colors = c("darkblue","lightgreen")) %>%
  layout(xaxis = list(title = "Engine Displacement in Litres"),
         yaxis = list(title = "Highway Miles per Gallon")) %>%
  colorbar(title = "# of Cylinders", y = 0.8)
# showing plot p4
p4
```
