---
title: "Stat 437 HW1"
author: 
     - Anthony Chelf (011743578)
header-includes:
   - \usepackage{bbm}
   - \usepackage{amssymb}
   - \usepackage{amsmath}
   - \usepackage{graphicx,float}
   - \usepackage{natbib}
output: 
  pdf_document:
    latex_engine: xelatex
fontsize: 11pt
---

```{r, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# General rule

Please show your work and submit your computer codes in order to get points. Providing correct answers without supporting details does not receive full credits. This HW covers:

- The basics of `dplyr`
- Creating scatter plot using `ggplot2`
- Elementary Visualizations (via ggplot2): density plot, histogram, boxplot, barplot, pie chart
- Advanced Visualizations via ggplot2: faceting, annotation

You DO NOT have to submit your HW answers using typesetting software. However, your answers must be legible for grading. Please upload your answers to the course space.

# Problem 1

Please refer to the NYC flight data `nycflights13` that has been discussed in the lecture notes and whose manual can be found at https://cran.r-project.org/web/packages/nycflights13/index.html. We will use `flights`, a tibble from `nycflights13`.

```{r}
# packages to be used
library(dplyr)
library(ggplot2)
library(nycflights13)

# peaking at the nycflights13 data
head(flights)
```


You are interested in looking into the average `arr_delay` for 6 different `month` 12, 1, 2, 6, 7 and 8, for 3 different `carrier` "UA", "AA" and "DL", and for `distance` that are greater than 700 miles, since you suspect that colder months and longer distances may result in longer average arrival delays. Note that you need to extract observations from `flights` and obtain the needed sample means for `arr_delay`, and that you are required to use `dplyr` for this purpose.

```{r}
filtered_flights <- filter(flights, month %in% c(12,1,2,6,7,8),
                           carrier %in% c('UA', 'AA', 'DL'), distance > 700)
```

The following tasks and questions are based on the extracted observations.

(1.a) In a single plot, create a density plot for `arr_delay` for each of the 6 months with `color` aesthetic designated by `month`. Note that you need to convert `month` into a factor in order to create the plot. What can you say about the average `arr_delay` across the 6 months?

```{r}
# converting month into a factor
month <- filtered_flights %>% mutate_if(is.integer,as.factor)
# generating the density plot
ggplot(month, aes(x = arr_delay, color = month)) +
   geom_density() + theme_grey() +ggtitle("Density Plot")
```
The average delay for each of our selected months are skewed to the left. It doesn't appear that there is much of a difference when looking at these specific months and that average delays tend to be minimal in nature. By minimal I mean close to 0.

(1.b) In a single plot, create a boxplot for `arr_delay` for each of the 3 carriers. What can you say about the average `arr_delay` for the 3 carriers?

```{r}
# creating a boxplot that shows the arr_delay for each of the 3 carriers
ggplot(filtered_flights, aes(x = carrier, y = arr_delay)) +
   geom_boxplot() + theme_grey() + ggtitle("Boxplot: arr_delay")
```
The average arr_delay for all 3 carriers appears to be close to 0.

(1.c) Create a pie chart for the 3 carriers where the percentages are the proportions of observations for each carrier and where percentages are superimposed on the sectors of the pie chart disc.

```{r}
# setting up the initial pie chart
flight.pc <- filtered_flights %>% group_by(carrier) %>%
   count() %>% ungroup() %>%
   mutate(percentage=n/sum(n)) %>%
   arrange(desc(carrier))
flight.pc$labels <- scales::percent(flight.pc$percentage)
# generating the pie chart with percentages added on top
ggplot(flight.pc) + geom_bar(aes(x="", y=percentage, fill=carrier),
                             stat="identity", width=1) +
   coord_polar("y", start=0) + ggtitle("Pie Chart: Carrier Observations") +
   geom_text(aes(x=1,y=cumsum(percentage)-percentage/2, label = labels))
```

(1.d) Plot `arr_delay` against `distance` with `facet_grid` designated by `month` and `carrier`.

```{r}
# generating a grid of plots and faceting by month and carrier
flights.plot <- ggplot(filtered_flights) +
   geom_point(mapping = aes(x=arr_delay, y=distance))
flights.plot + facet_grid(month ~ carrier)
```

(1.e) For each feasible combination of values of `month` and `carrier`, compute the sample average of `arr_delay` and save them into the variable `mean_arr_delay`, and compute the sample average of `distance` and save these averages into the variable `mean_distance`. Plot `month` against `mean_arr_delay` with `shape` designated by `carrier` and `color` by `mean_distance`, and plot `month` against `mean_arr_delay` with `shape` designated by `carrier` and `color` by `mean_distance` and annotate each point by its associated `carrier` name. 

```{r}
# utilizes piping to find combination of values of month/carrier
# and computes averages of delay and distance
temp <- filtered_flights %>% select(month, carrier, arr_delay, distance) %>%
   filter(month %in% c(1, 2, 6, 7, 8, 12), carrier %in% c("AA", "UA", "DL"))
temp <- na.omit(temp)
temp <- temp %>% group_by(month, carrier) %>%
   summarise(mean_arr_delay = mean(arr_delay),
             mean_distance = mean(distance), .groups="keep") %>% as.data.frame()
#generates the first plot with no annotation
plot <- ggplot(temp, aes(x = month, y = mean_arr_delay)) +
   geom_point(aes(shape = carrier, color = mean_distance)) +
   ggtitle("Plot 1")
plot
# generates the second plot with annotation
plot + geom_text(aes(label = carrier), size = 2, hjust = 1.2) +
   ggtitle("Plot 2")
```

# Problem 2

Please refer to the data set `mpg` that is available from the `ggplot2` package. Plot `displ` against `hwy` with faceting by `drv` and `cyl`, `color` designated by `class`, and `shape` by `trans`. This illustrates visualization with 4 factors.

```{r}
# peaking at mpg data
head(mpg)
```

```{r}
# generating plots for displ vs hwy and faceting by drv and cyl
plot.mpg <- ggplot(mpg, aes(x=displ, y=hwy)) + facet_grid(drv~cyl) +
   geom_point(aes(color=class, shape=trans))
plot.mpg
```

```{r}
# adding scale_shape_manual to handle the shape palette limitation of 6
plot.mpg <- ggplot(mpg, aes(x=displ, y=hwy)) + facet_grid(drv~cyl) +
   geom_point(aes(color=class, shape=trans)) + scale_shape_manual(values =
1:length(unique(mpg$trans)))
plot.mpg
```

