---
title: "Quiz 2"
author: "Anthony Chelf"
date: "10/11/2021"
output: 
  pdf_document:
    latex_engine: xelatex
---

# Question 1a
Simulate a data set as follows

```{r}
set.seed(1)
```

```{r}
n = 300; p = 200; s = 5
```

```{r}
x = matrix(rnorm(n*p), n, p)
b = c(rep(1, s), rep(0, p-s))
y = 1+x%*%b+rnorm(n)
```

# Question 1b
Define a grid L = {0,...,2} of 100 numbers between 0 and 2, this grid shall serve as potential values for the regularizer lambda

```{r}
L = seq(0, 2, length.out = 100)
L
```

# Question 1c
Use the function glmnet() to obtain lasso estimates for each value in the grid that you define in part(b). Using the function coef(), extract the estimated coefficient vector when lambda = L[10], i.e., the 10th value in the grid L.

```{r}
library(glmnet)
library(MASS)
```
```{r}
lasso.mod = glmnet(x, y, lambda = L, alpha = 1)
summary(lasso.mod)
```
```{r}
hbeta=coef(lasso.mod, s = 10)
as.numeric(hbeta)
```
# Question 1d
For each value of lambda in the grid L of part(b), compute the mean squared error on the entire dataset. (Use a for() loop for this purpose). This will provide a vector of 100 values of mse, one for each value of lambda. Plot lambda vs mse. What do you observe?

I am unsure of how to continue from here after reviewing the lectures, labs, and relevant information in ISLR.

```{r}

```

# Question 1e
Using a for() loop. compute the cross validation error, for each value of lambda in the grid L, under a k = 5
fold cross validation setup. (you will need to repeatedly divide the data into testing and training, this
will require another for() loop).
```{r}

```

# Question 1f
Compile all your code into a custom function with input arguments x, y, k and a grid L. This function
should output the following results, (i) the best fit model with a k-fold cross validation, (ii) the vector
of cross validation errors (one for each value of lambda), (iii) the grid L used for cross validation, (iv) the
value of lambda at which the best fit model is obtained.
```{r}


```

# Question 1g
Finally, use the function you make in Part (f) with k = 5, then extract the vector of cross validation
errors (say, CVV) and the grid L that is used. Make a plot of L vs. CVV.
```{r}

```

